---
title: "Figure 7 Colon CA Extract metrics"
author: "Yuzhou Feng"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(SPIAT)
library(ggplot2)
library("RColorBrewer")
library(dplyr)
library(reshape2)

data <- read.delim("data/CRC_clusters_neighborhoods_markers.csv", sep=",")
```

```{r}
# define a function to extract metrics
run_basic_metrics <- function(local_patient, phenotype_map){
  local_patient_pheno <- local_patient$ClusterName
  local_patient_intensity <- t(local_patient[,c(13:61, 70:78)])
  colnames(local_patient_intensity) <- local_patient$CellID
  
  rownames(local_patient_intensity) <-
      c("CD44","FOXP3","CD8","p53","GATA3","CD45","T.bet","beta.catenin","HLA.DR","PD.L1",
        "Ki67","CD45RA","CD4","CD21","MUC.1","CD30","CD2","Vimentin","CD20","LAG.3",
        "Na.K.ATPase","CD5","IDO.1","Cytokeratin","CD11b","CD56","aSMA","BCL.2",
        "CD25","CD11c","PD.1","Granzyme.B","EGFR","VISTA","CD15","ICOS","Synaptophysin",
        "GFAP","CD7","CD3","Chromogranin.A","CD163","CD45RO","CD68","CD31","Podoplanin",
        "CD34","CD38","CD138","HOECHST1","CDX2","Collagen.IV","CD194","MMP9",
        "CD71","CD57","MMP12","DRAQ5")
  
  local_patient_image <- format_image_to_spe(format = "general", intensity_matrix = local_patient_intensity,
                                             phenotypes = local_patient_pheno, coord_x = local_patient$X.X, coord_y = local_patient$Y.Y)
  
  local_patient_image <- define_celltypes(local_patient_image,
                                          category_colname = "Phenotype",
                                          names = phenotypes_map$Type,
                                          categories = phenotypes_map$Name,
                                          new_colname = "Cell.Type2") 
  
  local_patient_image <- define_celltypes(local_patient_image,
                                          category_colname = "Phenotype",
                                          names = phenotypes_map$Name,
                                          categories = phenotypes_map$Name,
                                          new_colname = "Cell.Type") 
  
  number_colors <- length(unique(local_patient_pheno))
  mycolors <- colorRampPalette(brewer.pal(8, "Set2"))(number_colors)
  
  plot_cell_categories(local_patient_image, unique(local_patient_pheno), mycolors, "Phenotype")

  p_cells <- calculate_cell_proportions(local_patient_image, feature_colname="Cell.Type")
  
  phenotypes <- phenotypes_map$Name
  
  NNI_df <- vector()
  for(pheno in phenotypes){
    NNI <- average_nearest_neighbor_index(local_patient_image, 
                                          reference_celltypes=pheno, 
                                          feature_colname="Cell.Type", p_val = 1e-2)
    NNI_df <- rbind(NNI_df, c(pheno, NNI$ANN_index, NNI$pattern, NNI$`p-value`))
  }
  
  colnames(NNI_df) <- c("Cell_type", "ANNI", "Pattern", "p_value")
  NNI_df <- as.data.frame(NNI_df)
  
  
  aveg_min_dist <- average_minimum_distance(local_patient_image)
  ##Return aveg_min_dist
  
  
  clusters_list <- identify_neighborhoods(
      local_patient_image, method = "hierarchical", min_neighborhood_size = 1,
    cell_types_of_interest = c("CD4+ T cells CD45RO+", "CD68+CD163+ macrophages", 
                               "plasma cells", "CD8+ T cells", "Tregs", "CD4+ T cells",
                               "CD11c+ DCs", "B cells", "CD11b+CD68+ macrophages",
                               "NK cells", "CD68+ macrophages GzmB+", "CD68+ macrophages",
                               "CD11b+ monocytes","CD4+ T cells GATA3+", 
                               "CD163+ macrophages", "CD3+ T cells"),
    radius = aveg_min_dist*3,
    feature_colname = "Cell.Type")
  
  T_cell_clusters_list <- identify_neighborhoods(
      local_patient_image, method = "hierarchical", min_neighborhood_size = 1,
      cell_types_of_interest = c("CD4+ T cells CD45RO+", "CD8+ T cells", 
                                 "Tregs", "CD4+ T cells", "CD4+ T cells GATA3+", 
                                 "CD3+ T cells"),
      radius = aveg_min_dist*3, feature_colname = "Cell.Type")
  
  macrophages_clusters_list <- identify_neighborhoods(
      local_patient_image, method = "hierarchical", min_neighborhood_size = 1,
      cell_types_of_interest = c("CD68+CD163+ macrophages", "CD11b+CD68+ macrophages",
                                 "CD68+ macrophages GzmB+",  "CD68+ macrophages", 
                                 "CD163+ macrophages"),
      radius = aveg_min_dist*3,  feature_colname = "Cell.Type")
  

  return(list(
    local_patient_image = local_patient_image,
    pro_cells = p_cells,
    NNI_df = NNI_df,
    ave_min_dist = aveg_min_dist,
    clusters = clusters_list,
    T_cell_clusters = T_cell_clusters_list,
    macrophages_clusters = macrophages_clusters_list
  ))
}

# phenotypes map
phenotypes_map <- data.frame(
    Name = c("granulocytes","vasculature","CD4+ T cells CD45RO+","tumor cells",
             "stroma","CD68+CD163+ macrophages","adipocytes","plasma cells",
             "CD8+ T cells","dirt","Tregs","CD4+ T cells", "CD11c+ DCs",
             "B cells","CD11b+CD68+ macrophages","smooth muscle", "undefined",
             "tumor cells / immune cells","immune cells","NK cells","nerves",
             "CD68+ macrophages GzmB+","CD68+ macrophages","lymphatics",
             "CD11b+ monocytes","CD4+ T cells GATA3+",
             "immune cells / vasculature","CD163+ macrophages","CD3+ T cells"),
    Type = c("Immune", "Vasculature", "Immune", "Tumor", "Stroma", "Immune", 
             "Adipocytes", "Immune", "Immune", "dirt",
             "Immune", "Immune", "Immune", "Immune", "Immune","Muscle", 
             "Undefined", "Undefined", "Immune", "Immune","Nerves", 
             "Immune", "Immune", "Lymphatics", "Immune", "Immune", 
             "Undefined","Immune", "Immune"))

data$patients_id <- paste("Patient", data$patients, data$File.Name, sep="_")

# extract metrics
basic_metrics_list <- list()

for(patient in unique(data$patients_id)){
  local_patient <- subset(data, patients_id == patient)
  local_patient$CellID <- paste("Cell", local_patient$CellID, sep="_")
  
  basic_metrics_list[[patient]] <- run_basic_metrics(local_patient, phenotype_map)
  save(basic_metrics_list, file="Objects/basic_metrics_list.Rdata")
}
